**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-06-15 submission/manuscript.Rmd
rmd updates

@@

```{r datasets}
meta <- '../data/raw/kws_metadata.tsv'
shared <- '../data/mothur/kws_final.an.shared'
tax <- '../data/mothur/kws_final.an.cons.taxonomy'
subsample <- '../data/mothur/kws_final.an.0.03.subsample.shared'

@@

tyc <- read.table("../data/mothur/kws_final.an.summary", sep = '\t', header = T, row.names=NULL)

@@

#use this code to assign phyla to each OTU in the shared file 
shared_phyla <- get_tax_level_shared(subsample, tax, 2)
phyla_met <- merge(meta_file, shared_phyla, by.x='group', by.y='row.names')

@@

#10 fold cross validation for L vs R lumen
iters <- 100
cv10f_aucs_lum <- c()
cv10f_all_resp_lum <- c()
cv10f_all_pred_lum <- c()
for(j in 1:iters){
  set.seed(j)
  sampling_lum <- sample(1:nrow(aucrf_data_LRlumen),nrow(aucrf_data_LRlumen),replace=F)
  cv10f_probs_lum <- rep(NA,39)
  for(i in seq(1,36,4)){
    train_lum <- aucrf_data_LRlumen[sampling_lum[-(i:(i+3))],]
    test_lum <- aucrf_data_LRlumen[sampling_lum[i:(i+3)],]
    set.seed(seed)
    temp_model_lum <- AUCRF(location~., data=train_lum, pdel=0.99, ntree=500)
    cv10f_probs_lum[sampling_lum[i:(i+3)]] <- predict(temp_model_lum$RFopt, test_lum, type='prob')[,2]
  }
  cv10f_roc_lum <- roc(aucrf_data_LRlumen$location~cv10f_probs_lum)
  cv10f_all_pred_lum <- c(cv10f_all_pred_lum, cv10f_probs_lum)
  cv10f_all_resp_lum <- c(cv10f_all_resp_lum, aucrf_data_LRlumen$location)
  cv10f_aucs_lum[j] <- cv10f_roc_lum$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc_lum <- roc(cv10f_all_resp_lum~cv10f_all_pred_lum)

@@

```{r important_otus}

>>>>

```{r knitr_settings, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

@@

```{r datasets}
meta <- '../data/raw/kws_metadata.tsv'
shared <- '../data/mothur/kws_final.an.shared'
tax <- '../data/mothur/kws_final.an.cons.taxonomy'
subsample <- read.table(file='../data/mothur/kws_final.an.0.03.subsample.shared', header=T)

@@

tyc <- read.table(file ="../data/mothur/kws_final.an.summary", sep = '\t', header = T, row.names=NULL)

@@

#use this code to assign phyla to each OTU in the shared file 
sub <- '../data/mothur/kws_final.an.0.03.subsample.shared'
shared_phyla <- get_tax_level_shared(sub, tax, 2)
phyla_met <- merge(meta_file, shared_phyla, by.x='group', by.y='row.names')

@@

#10 fold cross validation for L vs R lumen
iters <- 100
cv10f_aucs_lum <- c()
cv10f_all_resp_lum <- c()
cv10f_all_pred_lum <- c()
for(j in 1:iters){
  set.seed(j)
  sampling_lum <- sample(1:nrow(aucrf_data_LRlumen),nrow(aucrf_data_LRlumen),replace=F)
  cv10f_probs_lum <- rep(NA,39)
  for(i in seq(1,36,4)){
    train_lum <- aucrf_data_LRlumen[sampling_lum[-(i:(i+3))],]
    test_lum <- aucrf_data_LRlumen[sampling_lum[i:(i+3)],]
    set.seed(seed)
    temp_model_lum <- AUCRF(location~., data=train_lum, pdel=0.99, ntree=500)
    cv10f_probs_lum[sampling_lum[i:(i+3)]] <- predict(temp_model_lum$RFopt, test_lum, type='prob')[,2]
  }
  cv10f_roc_lum <- roc(aucrf_data_LRlumen$location~cv10f_probs_lum)
  cv10f_all_pred_lum <- c(cv10f_all_pred_lum, cv10f_probs_lum)
  cv10f_all_resp_lum <- c(cv10f_all_resp_lum, aucrf_data_LRlumen$location)
  cv10f_aucs_lum[j] <- cv10f_roc_lum$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc_lum <- roc(cv10f_all_resp_lum~cv10f_all_pred_lum)
```
```{r important_otus}
#just LB vs LS 
left_otu_feat <- colnames(aucrf_data_left_bs[2:6])
otu_taxa_left <- get_tax(1, left_otu_feat, tax_file)
#Abundance stripchart or most predictive otus
ls_abunds <- shared_meta[shared_meta$location=='LS', left_otu_feat]/10000 + 1e-4
lb_abunds <- shared_meta[shared_meta$location=='LB', left_otu_feat]/10000 + 1e-4

right_otu_feat <- colnames(aucrf_data_right_bs[2:6])
otu_taxa_right <- get_tax(1, right_otu_feat, tax_file)
#Abundance stripchart or most predictive otus
rs_abunds <- shared_meta[shared_meta$location=='RS', right_otu_feat]/10000 + 1e-4
rb_abunds <- shared_meta[shared_meta$location=='RB', right_otu_feat]/10000 + 1e-4

#Lb vs Rb
LRbowel_otu_feat <- colnames(aucrf_data_LRbowel[2:6])
otu_taxa_LRbowel <- get_tax(1, LRbowel_otu_feat, tax_file)
#Abundance stripchart or most predictive otus 
lb_abunds <- shared_meta[shared_meta$location=='LB', LRbowel_otu_feat]/10000 + 1e-4
rblb_abunds <- shared_meta[shared_meta$location=='RB', LRbowel_otu_feat]/10000 + 1e-4

#LS vs RS
LRlumen_otu_feat <- colnames(aucrf_data_LRlumen[2:6])
otu_taxa_LRlumen <- get_tax(1, LRlumen_otu_feat, tax_file)
#Abundance stripchart or most predictive otus 
lsrs_abunds <- shared_meta[shared_meta$location=='LS', LRlumen_otu_feat]/10000 + 1e-4
rsls_abunds <- shared_meta[shared_meta$location=='RS', LRlumen_otu_feat]/10000 + 1e-4

@@

```{r oncogenic_otus}
subsampled_meta <- merge(meta_file, subsample, by.x='group', by.y='row.names')
fuso179 <- subsampled_meta[, colnames(subsampled_meta) %in% c("group", "patient", "location", "Otu00179")]

fuso179[,5] <- (fuso179[,4]/subsampled_to)*100
names(fuso179)[5] <- "Otu00179_relAbund"

p_152 <-subsampled_meta[, colnames(shared_meta) %in% c("Group", "patient", "location", "Otu00152")]

p_152[,4] <- (p_152[,3]/subsampled_to) *100
names(p_152)[4] <- "Otu152_abund"
```
